---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
download.file("http://d396qusza40orc.cloudfront.net/repdata/data/activity.zip", "activity.zip")
unzip("activity.zip", exdir = "activity")
activity<-read.csv("activity/activity.csv")

temp2 <- mapply(function(x, y) paste0(rep(x, y), collapse = ""), 0, 4 - nchar(activity$interval))
activity$interval <- paste0(temp2, activity$interval)

```

## What is mean total number of steps taken per day?

```{r, echo=TRUE}
library(plyr)

steps_select<-subset(activity,select=c(date,steps))
steps_clean<-na.omit(steps_select)
steps_total<-ddply(steps_clean,"date", summarise, total=sum(steps))
hist(steps_total$total,breaks=30,
     main='Total no. of steps taken each day',
     xlab='Sum of steps'
     )

steps_mean<-mean(steps_total$total)
steps_median<-median(steps_total$total)
steps_mean
steps_median

```

```

## What is the average daily activity pattern?

```{r, echo=TRUE}

library(ggplot2)
library(reshape2)
library(scales)

steps_select2<-subset(activity,select=c(interval,steps))
steps_clean2<-na.omit(steps_select2)
steps_mean2<-ddply(steps_clean2,"interval", summarise, mean=mean(steps))
ggplot(steps_mean2,aes(x=strptime(interval, format="%H%M"), y=mean, group=1))+geom_line()+scale_x_datetime(labels=date_format("%H:%M"))+xlab("Time")

steps_max<-subset(steps_mean2,mean==max(mean))
steps_max

#Time where the no. of steps taken is at max on average is at 8:35am.

```


## Imputing missing values

```{r, echo=TRUE}

#Imputing missing values with the mean for that particular day.

sum(is.na(activity$steps))
set_to_zero<-activity
set_to_zero[is.na(set_to_zero)]<-0
steps_mean_day<-ddply(set_to_zero,"date", summarise, mean=mean(steps))
combined<-merge(steps_mean_day,activity,by="date")
combined$new <- ifelse(is.na(combined$steps), combined$mean, combined$steps)
steps_total_new<-ddply(combined,"date", summarise, total=sum(new))
hist(steps_total_new$total,breaks=30,
    main='Total no. of steps taken each day_imputed missing',
    xlab='Sum of steps')

steps_mean_new<-mean(combined$new)
steps_median_new<-median(combined$new)
steps_mean_new
steps_median_new

# The histogram now reflects that in fact, one of the highest frequencies in a day is where one does not take any steps at certain times of the day. This also brought down the average no. of steps taken per day.

```

## Are there differences in activity patterns between weekdays and weekends?

```{r, echo=TRUE}

combined$days<-weekdays(as.Date(combined$date))
combined$days_label<-ifelse(combined$days=="Saturday"|combined$days=="Sunday",
                            "Weekend","Weekday")
steps_combined<-ddply(combined,c("days_label","interval"), summarise, mean=mean(new))
lc<-ggplot(steps_combined,aes(x=strptime(interval, format="%H%M"), y=mean, group=1))+geom_line()+scale_x_datetime(labels=date_format("%H:%M"))+xlab("Time")

lc+facet_grid(days_label~.)

#Generally less steps were taken on weekend than on weekdays and steps taken on weekends do not vary as much as on weekedays.

```


